Primal-dual algorithm for computing Nash equilibria in sequential games with imcomplete information

### 2015-01-22 nips2015.tex
more refs

\section{Introduction}
\label{sec:intro}
A game-theoretic aproach to playing games strategically optimally consists in computing Nash-equilibria (infact, approximations thereof) offline, and playing one's part of the equilibrium online. This technique is the driving-force behind solution concepts like CFR \cite{lanctot2009monte}, $\text{CFR}^{+}$, and other variants, which have recently found profound success in Poker: equilibria are computed offline (with tremendous computation resources, and the component optimal strategies are played on line by a bot, without any a posteriori computation). However, solving for equilibria remains a mathematical and computational challenge, especially in sequential games with imperfect information. Under the assumption of \textit{perfect recall}, \cite{von1996efficient} and co-workers have developed the \textit{sequence-form} representation, whose size is linear in the size of the game tree, in contrast with the \textit{strategic-form} representation which is exponential. The goal of this paper is to develop a fast and simple primal-dual algorithm for computing  Nash $\epsilon$-equilibria for such games, using this representation.

\subsection{Background}
In \cite{von1996efficient, vonequilibrium}, it was established that for sequential two-person zero-sum games with imcomplete information and perfect recall, there exist sparse matrices $A \in \mathbb{R}^{n_1 \times n_2}, E_1 \in \mathbb{R}^{l_1 \times n_1}, E_2 \in \mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in \mathbb{R}^{l_1}, e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$, $l_1$, and $l_2$ are all linear in the size of the game tree (number of states in the game) and such that Nash equilibria correspond to pairs $(x, y)$ of \textit{realization plans} which solve the primal LCP (Linear Constrained Programming) problem
\begin{equation}
  \begin{aligned}
    & \underset{y,p}{\text{minimize}}
    & & p^Te_1 \\
    & \text{subject to}
    & & -Ay + E_1^Tp \geq 0,\\
    &&& E_2y = e_2,\\
    &&& y \ge 0
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

@@

It was also shown (for example, Theorem 3.14 of \cite{vonequilibrium}) that a $(x, y)$ is a solution to the LCP problems \eqref{eq:primal_pb} and \eqref{eq:dual_pb} iff there exist vectors $p$ and $q$ such that
\begin{equation}
  \begin{split}
    E_1x = e_1,\text{ }x \geq 0,\hspace{1em} E_2y = e_2,\text{ }y \geq 0\\
    -Ay + E_1^Tp \geq 0,\hspace{1em}-A^Tx + E_2^Tq \leq 0
    \end{split}
\end{equation}

@@

\section{Related work}
\label{sec:related_work}
\subsection{Interior-point methods}
...

\subsection{Smoothing methods}
In \cite{hoda2010smoothing}, a nested iterative procedure using the Excessive Gap Technique \cite{nesterov2005excessive} was used to to solve the LCP problems \eqref{eq:primal_pb} and \eqref{eq:dual_pb} in the following equivalent saddle-point form

@@

\subsection{CounterFactual Regret minimization}
More recently, the primal-dual algorithm was applied to matrix games (on simplexes) in \cite{chambolle2014ergodic}. Sampling techniques like the CFR (CounterFactual Regret minimization) and its many variants \cite{MartinZinkevichNIPS2007, lanctot2009monte} are also becoming popular...

@@

\subsection{Unconstrained minimax reformulation of the problem}
It should be noted that the difficulty of applying the general primal-dual Algorithm 39 proposed in \cite{chambolle2010} directly to problem \eqref{eq:primal_pb} lies in the difficulty of effectively computing\footnote{However, this computation can be done iteratively using the algorithm in proposition 4.2 of \cite{combettes2010dualization}, for example.} the euclidean projections $\Pi_{Q_k}$. Theorem \ref{thm:pd} provides a trick to effectively evade this difficulty.

@@

Applying Algorithm 39 of \cite{chambolle2010} to problem \eqref{eq:unconstrained_pb}, we obtain Algorithm \ref{Tab:algo_simplified} for solving the original problem \eqref{eq:primal_pb}. The $\mathcal{O}(1/\epsilon)$ convergence rate of Algorothm \ref{Tab:algo_simplified} hails directly from Theorem 1 of \cite{chambolle2010}.

Note that the convergence rate of $\mathcal{O}(1/\epsilon)$ cannot be improved in the framework of \cite{chambolle2010}, since neither $G_1$ nor $G_2$ is \textit{uniformly convex}.
However, it's possible to do over-relaxation and inertial iterates...

>>>>

\section{Introduction}
\label{sec:intro}
A game-theoretic aproach to playing games strategically optimally consists in computing Nash-equilibria (infact, approximations thereof) offline, and playing one's part of the equilibrium online. This technique is the driving-force behind solution concepts like CFR \cite{lanctot2009monte}, $\text{CFR}^{+}$, and other variants, which have recently found profound success in Poker: equilibria are computed offline (with tremendous computation resources, and the component optimal strategies are played on line by a bot, without any a posteriori computation). However, solving for equilibria remains a mathematical and computational challenge, especially in sequential games with imperfect information. This paper proposes an algorithm for solving for such equilibria approximately, in a sense which will be made clear shortly.

\subsection{Context}
The sequence-form for two-person zero-sum games was introduced in \cite{koller1992complexity}, and the theory was further developed in \cite{von1996efficient, vonequilibrium}, where it was established that for sequential two-person zero-sum games with imcomplete information and perfect recall, there exist sparse matrices $A \in \mathbb{R}^{n_1 \times n_2}, E_1 \in \mathbb{R}^{l_1 \times n_1}, E_2 \in \mathbb{R}^{l_2 \times n_2}$, and vectors $e_1 \in \mathbb{R}^{l_1}, e_2 \in \mathbb{R}^{l_2}$ such that $n_1$, $n_2$, $l_1$, and $l_2$ are all linear in the size of the game tree (number of states in the game) and such that Nash equilibria correspond to pairs $(x, y)$ of \textit{realization plans} which solve the primal LCP (Linear Constrained Programming) problem
\begin{equation}
  \begin{aligned}
    & \underset{y,p}{\text{minimize}}
    & & p^Te_1 \\
    & \text{subject to}
    & & -Ay + E_1^Tp \geq 0,\\
    &&& E_2y = e_2,\\
    &&& y \ge 0
  \end{aligned}
  \label{eq:primal_pb}
\end{equation}

@@

It was also shown (see Theorem 3.14 of \cite{vonequilibrium}) that a $(x, y)$ is a solution to the LCP problems \eqref{eq:primal_pb} and \eqref{eq:dual_pb} iff there exist vectors $p$ and $q$ such that
\begin{equation}
  \begin{split}
    E_1x = e_1,\text{ }x \geq 0,\hspace{1em} E_2y = e_2,\text{ }y \geq 0\\
    -Ay + E_1^Tp \geq 0,\hspace{1em}-A^Tx + E_2^Tq \leq 0
    \end{split}
\end{equation}

@@

\section{Related work}
In \cite{hoda2010smoothing}, a nested iterative procedure using the Excessive Gap Technique \cite{nesterov2005excessive} was used to to solve the LCP problems \eqref{eq:primal_pb} and \eqref{eq:dual_pb} in the following equivalent saddle-point form

@@

Sampling techniques like the CFR (CounterFactual Regret minimization) and its many variants \cite{MartinZinkevichNIPS2007, lanctot2009monte, Bowling09012015} are also becoming popular.

More recently, \cite{chambolle2014ergodic} employs the primal-dual algorithm was applied to matrix games on simplexes. It should be stressed that such matrix games are considerably simpler than the games considered here. Indeed, the authors in \cite{chambolle2014ergodic} use the fact that computing the euclidean projection of a point unto a simplex can be done in linear time \cite{duchi2008efficient}, in contrast to the polytopes $Q_k$ defined in \eqref{eq:polytope} for which no such efficient algorithm is known nor is likely to exist. Note that computing the euclidean projection of a point unto the polytope $Q_k$ can still be done iteratively using the algorithm in proposition 4.2 of \cite{combettes2010dualization}, for example. However, the authors didn't provide a convergence rate for their algorithm, and as with any nested iterative scheme, one would have to solve this sub-problem with finer and finer precision.

@@

\subsection{Unconstrained minimax reformulation of the problem}
It should be noted that the difficulty of applying the general primal-dual Algorithm 39 proposed in \cite{chambolle2010} directly to problem \eqref{eq:primal_pb} lies in the difficulty of effectively computing the euclidean projections $\Pi_{Q_k}$. Theorem \ref{thm:pd} provides a trick to effectively evade this difficulty.

@@

Applying Algorithm 39 of \cite{chambolle2010} to problem \eqref{eq:unconstrained_pb}, we obtain Algorithm \ref{Tab:algo_simplified} for solving the original problem \eqref{eq:primal_pb}. The $\mathcal{O}(1/\epsilon)$ convergence rate of Algorothm \ref{Tab:algo_simplified} hails directly from Theorem 1 of \cite{chambolle2010}. Note however that this convergence rate cannot be improved in the framework of \cite{chambolle2010}, since neither $G_1$ nor $G_2$ is \textit{uniformly convex}.
